The Pair Distribution Function of the Inhomogeneous Hard-Sphere Fluid

### 2013-11-21 papers/pair-correlation/paper.tex
make plot-ghs2.py generate numbers to include (or input) directly into *.cpp (or paper.tex) Tested on collins. Took 0:30.35.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
\end{figure}
\subsection{Separable fit to the radial distribution function}

@@

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these two constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi h_\sigma (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent how the on the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions of less
than 0.4 is \fixme{get a number}.

\begin{table}
  \begin{tabular}{cc}
    $\kappaa$ & 1 \\
    $\kappab$ & 1 \\
    $\kappac$ & 1 \\
  \end{tabular}
  \caption{Fit parameters \fixme{... put them here and format this nicely}}
  \label{tab:fit-parameters}
\end{table}

@@

\appendix

\section{Properties of $g_\sigma$}
We begin with some definitions of basic quantities, looking at the
homogeneous hard-sphere fluid.  I am ignoring here the distinction
between $N$ and $N-1$, which is unimportant in the thermodynamic limit
$N\gg 1$.
\begin{align}
  Z(\sigma) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
  \\
  n^{(2)}(\rr_1,\rr_2) &= \frac{N^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
\end{align}
\begin{align}
  g^{(2)}(\rr_1,\rr_2) &=
  \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
  \\
  &= \frac{V^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j}
  \Theta(r_{ij}-\sigma)
  \label{eq:g2-integral-defn}
\end{align}

At this point we consider the effect of increasing the diameter
$\sigma$ of all our spheres.  This will reduce the number of available
states, thus reducing the partition function $Z$, which we will
ultimately relate to a change in free energy.
\begin{align}
  Z(\sigma+\delta) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma-\delta)
  \\
  &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)(1-\Theta(\sigma+\delta-r_{ij}))
\end{align}
I have written $Z(\sigma+\delta)$ in such a way that it is clear that
it is equal to $Z(\sigma)$ plus a small quantity, provided $\delta$ is
small.  We will now investigate that small quantity, which we will
write as a power series in $\delta$.  We begin by expanding the
product of $(1-\Theta(\sigma+\delta-r_{ij}))$:
\begin{multline}
    \prod_{i<j}(1-\Theta(\sigma+\delta-r_{ij})) =
    1 - \sum_{i<j} \Theta(\sigma+\delta-r_{ij}) \\
    + \frac{1}{2}\mathop{\mathop{\sum_{i<j}}_{k<l}}_{kl \ne ij}
    \Theta(\sigma+\delta-r_{ij}) \Theta(\sigma+\delta-r_{kl})
    \\ + \cdots.
\end{multline}
Given this expansion, we can readily expand the partition function:
\begin{multline}
  Z(\sigma+\delta) = Z(\sigma) \\ -
  \sum_{k<l} \iiiint \Theta(\sigma+\delta-r_{kl}) \prod_{i<j}
  \Theta(r_{ij}-\sigma) \\
  + \frac{1}{2}\mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
  \iiiint \Theta(\sigma+\delta-r_{kl}) \Theta(\sigma+\delta-r_{mn}) \prod_{i<j}
  \Theta(r_{ij}-\sigma) \\
  + \cdots.\label{eq:z-power-series}
\end{multline}
Our challenge now is to manipulate this into forms involving $g(r)$,
which we can then simplify using a Taylor expansion around
$r=\sigma$.  We will begin by taking advantage of the fact that the
summations in Eq.~\ref{eq:z-power-series} are mostly of identical
terms, since the integrals are over all coordinates.
\begin{multline}
  Z(\sigma)-Z(\sigma+\delta) =
  \frac{N^2}{2}\iiiint \Theta(\sigma+\delta-r_{12})
  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
  - \frac{N^4}{8}\iiiint
  \Theta(\sigma+\delta-r_{12})
  \Theta(\sigma+\delta-r_{34})
  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
  + \cdots \label{eq:z-expansion}
\end{multline}
At this point we are ready to connect with $g(r)$, which we can do by
considering an integral of the pair distribution function, from
Eq.~\ref{eq:g2-integral-defn}.
\begin{multline}
  \int_{\sigma}^{\sigma+\delta} g(r) d\rr\\
  \shoveleft{\quad= \int g^{(2)}(\rr_1,\rr_2)
  \Theta(\sigma+\delta-r_{12})d\rr_2} \\
  \shoveleft{\quad= \frac{1}{V}\int g^{(2)}(\rr_1,\rr_2)
  \Theta(\sigma+\delta-r_{12})d\rr_1d\rr_2} \\
  \shoveleft{\quad= \frac{V}{Z}\iiiint d\rr_1\cdots d\rr_N
  \Theta(\sigma+\delta-r_{12})}\\
  \prod_{i<j} \Theta(r_{ij}-\sigma)
  \label{eq:g-shell-integral}
\end{multline}
Looking back at Eq.~\ref{eq:z-expansion}, we can see that the first
term is easily related to Eq.~\ref{eq:g-shell-integral}.  The second
term has two shell $\delta$ functions, and is thus a bit trickier.
Fortunately, in a very large system, $\rr_3$ and $\rr_4$ are
vanishingly seldom near $\rr_1$ or $\rr_2$, so we can write the second
term as a product.
\begin{multline}
  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)}\\
  \shoveleft{\quad= \frac{nN}{2} \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
  - \frac{n^2N^2}{8} \left(\int_{\sigma}^{\sigma+\delta} g(r)
  d\rr\right)^2
  + \cdots
  \label{eq:zdiff-g}
\end{multline}
Now let us look at how this integral relates to $g(r)$ evaluated at
contact.  I'll skip a few steps doing this integral...
\begin{equation}
   \int_{\sigma}^{\sigma+\delta} g(r)
  d\rr = 4\pi\sigma^2\left(g_\sigma\delta + \left(\tfrac12g_\sigma' +
  \frac{g_\sigma}{\sigma}\right)\delta^2 + \cdots \right)
\end{equation}
Putting this together with the above we find
\begin{multline}
  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
  2\pi\sigma^2 n N g_\sigma \delta +
  \pi\sigma n N  \left(g_\sigma'\sigma + 2g_\sigma\right)
  \delta^2 \\
  - 2\pi^2\sigma^4n^2N^2g_\sigma^2\delta^2 + \mathcal{O}(\delta^3)
\end{multline}
But by considering a Taylor expansion, we can also see that
\begin{multline}
  Z(\sigma)-Z(\sigma+\delta) =
  -\frac{\partial Z}{\partial \sigma} \delta
  -\frac12 \frac{\partial^2 Z}{\partial \sigma^2} \delta^2  + \mathcal{O}(\delta^3)
\end{multline}
So that
\begin{multline}
  \frac{\partial Z}{\partial \sigma} =
  -2ZnN\pi\sigma^2g_\sigma\\
  \frac{\partial^2 Z}{\partial \sigma^2} =
  -2ZnN\pi\sigma\left(g_\sigma'\sigma + 2g_\sigma \right)
  +4Zn^2N^2\pi^2\sigma^4{g_\sigma}^2
\end{multline}
from which we can extract relationships for the derivatives of $Z$
with respect to $\sigma$.  Which if we knew what those derivatives
were, would tell us about $g_\sigma'$ and $g_\sigma$, which is what we
are after here.  We find this out by considering the Helmholtz free
energy.
\begin{equation}
  F = -kT \ln Z
\end{equation}
The derivatives are then given by
\begin{equation}
  \frac{\partial F}{\partial \sigma} = -\frac{kT}{Z} \frac{\partial
    Z}{\partial \sigma}
\end{equation}
and
\begin{equation}
  \frac{\partial^2 F}{\partial \sigma^2} =
  \frac{1}{kT}\left(\frac{\partial F}{\partial \sigma}\right)^2
  -\frac{kT}{Z} \frac{\partial^2
    Z}{\partial \sigma^2}.
\end{equation}
With a bunch more work, we find out that
\begin{equation}
  g_\sigma' = \frac{\frac{\partial^2 F}{\partial
      \sigma^2}}{2\pi\sigma^2nNkT}
  - \frac{2g_\sigma}{\sigma}
\end{equation}
except that there is a mistake in here.  I can tell because $F$ is
extensive and $g_\sigma'$ is not.  This tracks back to
Eq.~\ref{eq:zdiff-g}, which is correct in this \LaTeX, but is wrong on
my scratch paper.

@@

\section{THE END OF THE PAPER}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Attard worked out the triplet correlation function, which is something
that we can access from the pair distribution function, if we consider
the pair correlation near a single hard sphere
solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al} wrote an
interesting paper using DFT and Monte Carlo to compute the
three-particle distribution function (same as the triplet correlation
function)\cite{gonzalez1999test}.  Their theory compares favorably
with the ``superposition approximation'', and we could probably do
something similar.

Plischke and Henderson worked out the pair correlation function near a
hard wall using Percus-Yevick, and compare with Monte Carlo
results\cite{plischke1986pair}.  They plot the pair correlation
function along interesting paths.  \fixme{Jeff read this paper for a
  while and does not understand what they do in oder to find the pair
  correlation function as compared to the radial distribution function}

Lado recently introduced a new and improved efficient algorithm for
implementing integral equation theory for inhomogeneous fluids, which
computes $g^{(2)}(\rr_1,\rr_2)$ among other
things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
use, and their example is limited to 1D systems.  And the scaling is
still (so far as I can tell) $O(N^2)$, since their improvement is in
the prefactor, by developing a nice basis set.

\section{Previous approaches}

Gloor \emph{et al} explain that little is known of the pair
distribution function of the hard-sphere fluid, and therefore use the
formula
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
\end{align}
They go on to further approximate the radial distribution function,
but we can just stop here.

Gloor et al \cite{gloor2007prediction} uses average density of two points:
\begin{align}
\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
\end{align}
and the uses bulk radial distribution for the reference system (hard
spheres) at this density, in order to approximate pair correlation.

Evans \cite{evans1992density}perscribes to use the same approximation but with an unspecified 'mean denisty':

\begin{align}
g^{(2)}(\rr_1,\rr_2) &\equiv g_r(\rho^m,\rr_{12})
\end{align}

Gross \cite{gross2009density} uses a hard chain reference fluid.  In
his functional he needs a pair correlation beween the different
segments of chains.  He approximates this function with the radial
distribution given by the Carnahan Starling at a density that is the
average density of the segment positions $\rr$ and $\rr'$:

\begin{align}
\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
\end{align}

Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but
maybe you (Dr. Roundy) should take a look at it.  They reference Tang
\cite{tang2008accurate} which I've mentioned below in the Appendix.
\fixme{DJR: Not same as Gloor, they do a WDA approximation.}

Sokolowski and Fischer \cite{sokolowski1992role} do the same as Gloor
et al except that instead of taking the exact densities
$\rho(\mathbf{r_1})$ and $\rho(\mathbf{r_2})$ to average he takes an
integral $\int_{V_{sphere}} \rho(\mathbf{r})d\mathbf{r}$ over a sphere
of some radius centered at each point.  It's these two averaged
densities that he takes the average of.

Fischer and Methfessel \cite{fischer1980born}, in a paper that is
cited by Sokolowski and Fischer (above) do the same thing (the
originate the idea of the spherical average density.  Their
Introduction has a very good discussion of the different averaging
methods to get to $g(\mathbf{r})$ tried up until then (paper is
written in 1980).

McCabe and Kiselev \cite{mccabe2004application} model a square well
potential.  One of their terms uses a radial distribution function
that once again is taken to be the bulk function evaluated at an
effective density.  They reference a source that calculates this
effective density (or filling fraction) to be $\eta_{effective} =
c_1\eta + c_2\eta^2 + c_3\eta^3$ where the coeffecients are gotten
from three coupled polynomial equations with $1,\lambda$, and
$\lambda^2$, where $\lambda$ is a measure of the model's potential
range.  They reference another source who calculates the effective
density in a similar (but more complicated) way.  I can look more at
these if you'd like.  They also use a different first order expansion
formula for the ditribution at contact in the chain term.

Toxvaerd \cite{toxvaerd1976hydrostatic}approximates the radial
distribution as the radial distribution of the bulk at the same
density.  Uses the first two terms in a tempurature expansion.  First
term is just Carnahan Starling derived, second is empirical.

\section{Pair distribution function with inhomogeneity}

The pair distribution function can be defined using the two particle
density as
\begin{align}
  g^{(2)}(\rr_1,\rr_2) &\equiv \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
\end{align}

We may also want to consider the \emph{triplet} distribution function
\begin{align}
  g^{(3)}(\rr_1,\rr_2,\rr_3) &\equiv \frac{n^{(3)}(\rr_1,\rr_2,\rr_3)}{n(\rr_1)n(\rr_2)n(\rr_3)}
\end{align}
The triplet distribution is relevant, because we can compute the
triplet distribution of the homogeneous system by considering the
\emph{pair} distribution of the inhomogeneous ``test-particle
scenario.''  The process is that we solve the inhomogeneous situation
in which we have one hard sphere located at the origin.  We solve for
the density around that hard sphere, and this gives us the pair
distribution function of the homogeneous system $g^{(2)}(0, \rr)$.
When we solve for the pair distribution $g^{(2)}(\rr_1,\rr_2)$ of the
inhomogeneous system, this will give us the \emph{triplet}
distribution function of the homogeneous system
$g^{(3)}(0,\rr_1,\rr_2)$.

\section{Goals of this paper (to be integrated into rest of paper)}

In this paper, we seek an efficient approximation for the pair
correlation function of an inhomogeneous hard-sphere fluid, where we
define efficient as an algorithm that scales as $O(N\log N)$ where $N$
scales as the volume of the system under consideration.
Integral equation theory (look this up) scales as $O(N^2)$, which can
be very expensive for large three-dimensional systems.

Fundamental-Measure Theory (FMT) is constructed using only single-point
convolutions (check terminology) such as
\begin{align}
  n_\alpha(\rr) &= \int n(\rr')w_\alpha(|\rr-\rr'|) d\rr'
\end{align}
These convolutions can be computed in Fourier space in $O(N\log N)$
time, with the result that the FMT free energy---and its
gradient---can be computed in $O(N\log N)$ time.

The pair distribution function is used in computing the first term in
the high-temperature perturbation expansion, which is used in SAFT
(for instance) to treat dispersion interactions.  This term is written
as
\begin{align}\label{eq:a1}
  a_1 &\propto \iint n(\rr)n(\rr') \Phi(|\rr-\rr'|)
  g^{(2)}(\rr,\rr')d\rr d\rr'
\end{align}
where $\Phi(r)$ is the pair potential.  If $g^{(2)}(\rr,\rr')$ could
be written as $g(|\rr-\rr'|)$, this term could be computed in $O(N\log
N)$ time.  However, this is not possible when the density is
inhomogeneous, because the pair distribution function depends on the
density.

In this paper, we consider a superposition approximation, in which
construct an approximate representation for the pair distribution
function, which allows the first term in the high temperature
expansion given in Equation~\ref{eq:a1} to be computed in $O(N\log N)$
time.  One input to this approach is our recently published expression
for an average of the value of the pair distribution function at
contact $g_\sigma(\rr)$, which itself can be computed in $O(N\log N)$
time.

\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, \eta(g_\sigma(\rr_1))) + g(|r_{12}|, \eta(g_\sigma(\rr_2)))}{2}
\end{align}
Bad:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = g\left(|r_{12}|, \frac{\eta(g_\sigma(\rr_1)) + \eta(g_\sigma(\rr_2))}{2}\right)
\end{align}

The general idea is to write
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \Theta(r_{12}-\sigma)
  \left(1 + \sum_{\alpha} \frac{f_{\alpha}(g_\sigma(\rr_1)) + f_{\alpha}(g_\sigma(\rr_2))}{2}g_{\alpha}(r_{12})\right)\label{eq:gpair-approx}
\end{align}
where $f_{\alpha}$ and $g_{\alpha}$ are a set of functions that
satisfy the following constraint:
\begin{align}\label{eq:definealpha}
  \sum_{\alpha} f_{\alpha}(g_\sigma(\eta)) g_{\alpha}(r) \approx g(r; \eta)
\end{align}
which ensures that our pair distribution function reduces
(approximately) to the radial distribution function in the homogeneous
limit.  Equation~\ref{eq:gpair-approx} is constructed such that
Equation~\ref{eq:a1} can be evaluated in $O(N\log N)$ time, since all
the required convolutions may be performed with a fixed kernel.

Our proposed pair distribution function has several good properties.
When evaluated at contact ($r_{12}=\sigma$), it gives the same good
result as our previously-published functional for the averaged pair
distribution function at contact, which is constructed to give an
accurate prediction for the contribution to Equation~\ref{eq:a1} from
spheres in contact.  In the homogeneous limit,
Equation~\ref{eq:gpqair-approx} should give a prediction that is as
accurate as the approximate equality in
Equation~\ref{eq:definealpha}.  In inhomogeneous distributions and at
intermediate distances, we expect our pair distribution function to be
least accurate, but one hopes that it will be reasonable, simply
because it is an average of two reasonable results.

\subsection*{Fundamental-Measure Theory}

We use the White Bear version of the Fundamental-Measure Theory~(FMT)
functional~\cite{roth2002whitebear}, which describes the excess free
energy of a hard-sphere fluid.  The White Bear functional reduces to
the Carnahan-Starling equation of state for homogeneous systems.  It
is written as an integral over all space of a local function of a set
of ``fundamental measures'' $n_\alpha(\rr)$, each of which is written
as a one-center convolution of the density.  The White Bear free
energy is thus
\begin{equation}
A_{\textit{HS}}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
\end{equation}
with integrands
\begin{align}
\Phi_1 &= -n_0 \ln\left( 1 - n_3\right) \label{eq:Phi1}\\
\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
  n_3 + (1-n_3)^2 \ln(1-n_3)
}{
  36\pi n_3^2\left( 1 - n_3 \right)^2
} , \label{eq:Phi3}
\end{align}
using the fundamental measures
\begin{align}
  n_3(\rr) &= \int n(\rr') \Theta(\sigma/2 -\left|\rr - \rr'\right|)
  d\rr' \label{eq:FMn3} \\
  n_2(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) d\rr' \\
  \mathbf{n}_{2V}(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
\end{align}
\begin{align}
  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{2\pi \sigma}, \quad
  n_1 &= \frac{n_2}{2\pi \sigma} , \quad
  n_0 = \frac{n_2}{\pi \sigma^2} \label{eq:FMrest}
\end{align}

\section{Theoretical Approaches}

The traditional approach for solving for the pair distribution
function of an inhomogeneous density distribution is to use an
integral equation theory...

\section{Comparison with simulation}\label{sec:comparison}

\appendix

\section*{Appendix}
\cite{jin2011perturbative} Combines hard sphere (white bear) with
attraction that is not mean field but rather uses two other
correltation expressions given in -\cite{tang2008accurate} and
\cite{hlushak2009direct}

Tang \cite{tang2008accurate}Derives the direct correltation for a SW
and Yukawa and LJ potentials by first getting the radial distribution
function through a method I don't really understand.  Composes fourier
transforms $c(k)$ and $h(k)$ of these complicated functions that
depend on filling fraction and the potential in an integral. Uses
first order mean spherical approximation.  This is just square
potential well around particle but the radial distribution function
takes only the first order in an expansion in... (high temp?).

\cite{hlushak2009direct} Cites \cite{tang2008accurate} (above) and
does largely the same thing but then adds a complication for $2 <
\lambda \leq 3$

\cite{levesque2012scalar} Uses direct correlation funtion that's
gotten from a radial distribution function of a homogeneous density.
The direct correlation itself is only dependent on distance.  The
$\rho_0$ that they use is the same $\rho_0$ that the used in the
reference HS term in their functional.  The use the Carnahan Starling
or the Perkis Yevick for these hard sphere free energies.  The other
terms involve $\delta\rho$ so the change is a variation from the
reference density.  They further modify it by looking at molecular
dynamic simulations of a certain model of water molecules and looking
at the oxygen-oxygen correlation.

%% G{\"o}tzelmann \emph{et al.} worked out the pair distribution function near a hard
%% wall by finding the direct correlation function by taking a functional
%% derivative of the free energy functional and then solving the
%% Ornstein-Zernicke equation~\cite{gotzelmann1996structure}.  They have plots like our pretty plots,
%% but with the fixed sphere one radius away from touching.
%% Interestingly, they plot the second bit as $z$, which enables them to
%% have just two axis values.  I think I prefer our ``total distance''
%% method, but there is a certain elegance to just plotting $z$ for the
%% curvy part as well as the straight-away part.  They found poor
%% quantitative agreement, particularly at contact.  They also studied
%% the pair correlation function in a hard slit~\cite{gotzelmann1997density}.

%% Integrating the Ornstein-Zernicke equation is challenging, requiring a
%% variational minimization itself~\cite{paul2003variational}.

%% Attard worked out the triplet correlation function, which is something
%% that we can access from the pair distribution function, if we consider
%% the pair correlation near a single hard sphere
%% solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al.} wrote an
%% interesting paper using DFT (test-particle approach of Percus) and
%% Monte Carlo to compute the three-particle distribution function (same
%% as the triplet correlation function)\cite{gonzalez1999test}.  Their
%% theory compares favorably with the ``superposition approximation'',
%% and we could probably do something similar.

%% Plischke and Henderson worked out the pair correlation function near a
%% hard wall using Percus-Yevick, and compare with Monte Carlo
%% results\cite{plischke1986pair}.  They plot the pair correlation
%% function along interesting paths.  \fixme{Jeff read this paper for a
%%   while and does not understand what they do in oder to find the pair
%%   correlation function as compared to the radial distribution function}

%% Lado recently introduced a new and improved efficient algorithm for
%% implementing integral equation theory for inhomogeneous fluids, which
%% computes $g^{(2)}(\rr_1,\rr_2)$ among other
%% things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
%% use, and their example is limited to 1D systems.  And the scaling is
%% still (so far as I can tell) $O(N^2)$, since their improvement is in
%% the prefactor, by developing a nice basis set.

%% Gloor \emph{et al} explain that little is known of the pair
%% distribution function of the hard-sphere fluid, and therefore use the
%% formula
%% \begin{align}
%%   g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
%% \end{align}
%% They go on to further approximate the radial distribution function,
%% but we can just stop here.

%% Gloor et al \cite{gloor2007prediction} uses average density of two points:
%% \begin{align}
%% \overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
%% \end{align}
%% and the uses bulk radial distribution for the reference system (hard
%% spheres) at this density, in order to approximate pair correlation.

%% \section{What has been done before?}
%% \subsection{Gloor simple method, problematic}
%% Gil-Villegas\cite{gil1997statistical} approximates the first order
%% inverse tenpurature Helmholtz Free energy term, which involves an
%% integral of a potential multiplied by a radial distribution function
%% $g^{HS}(r,\eta)$, by first taking the distribution function out of the
%% integral and evaluating it at the correct mean distance $\zeta$,
%% according to the mean value theorem.  They then approximate the pair
%% correlation function $g^{HS}(\zeta,\eta)$ with the correlation at
%% contact and at a effective $\eta_{eff}$.  The $\eta_{eff}$ is found by
%% fitting a general function form to correct energy solutions $a_1$ with
%% depenedent variable $\eta$ and

%% \subsection{Fischer accurate but slow}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
\end{figure}

\input{figs/fit-parameters}

\subsection{Separable fit to the radial distribution function}

@@

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these two constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi h_\sigma (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent how the on the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions of less
than 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr$.

\begin{table}
  \begin{tabular}{cc}
    $\kappaa$ & \kappazero \\
    $\kappab$ & \kappaone \\
    $\kappac$ & \kappatwo
  \end{tabular}
  \caption{Fit parameters \fixme{... put them here and format this nicely}}
  \label{tab:fit-parameters}
\end{table}

@@

@@
